ms_figures

### 2023-01-06 ms_figures.Rmd
text and figure edits for full draft

library(tidyverse)
library(countrycode)
library(circlize)
library(here)
library(ggpubr)
library(kableExtra)
library(DBI)
library(janitor)
library(viridis)
library(broom)
library(exploreARTIS)
library(vegan)
```

@@

artis <- artis %>%
  select(-record_id) %>%
  # Remove 2020 for now
  filter(year < 2020)

prod <- dbGetQuery(con, "SELECT * FROM production")
prod <- prod %>%
  select(-record_id) %>%
  # Remove 2020 for now
  filter(year < 2020)

@@

@@

# Set directories 
model_inputs_dir <- "/Volumes/jgephart/ARTIS/Outputs/model_inputs_20220926"
outdir <- "/Volumes/jgephart/ARTIS/Outputs/ms_seafood_globalization"

@@

```{r load_drive_data}
code_max_resolved_taxa <- read.csv("/Volumes/jgephart/ARTIS/Outputs/clean_metadata/code_max_resolved_taxa.csv") %>%
              mutate(hs6 = as.character(hs6))

sciname_metadata <- read.csv("/Volumes/jgephart/ARTIS/Outputs/clean_metadata/sciname_metadata.csv")

@@

```{r join_data}
# Load cleaned FAO production data
prod <- prod %>% 
  select("year", "iso3c", "sciname", 
         "method", "habitat", 
         "production_t" = "live_weight_t") %>%
  group_by(year, iso3c, sciname, method, habitat) %>%
  summarise(production_t = sum(production_t, na.rm = TRUE)) %>%
  ungroup()

@@

# Update artis data to lowest resolution by comparing code taxa_level and prod taxa_level
artis <- artis %>%
  # FIX IT: Change metadata to HSXX format and then remove this line
  mutate(hs_version = as.integer(parse_number(hs_version))) %>%
  left_join(code_max_resolved_taxa,
            by = c("hs_version", "hs6", "sciname")) %>%
  left_join(sciname_metadata, 
            by = "sciname") %>%
  # Leave any missing sciname_hs_modified as original sciname
  mutate(sciname_hs_modified = case_when(
    is.na(sciname_hs_modified) ~ sciname,
    TRUE ~ sciname_hs_modified
  )) %>%
  rename(environment = habitat) %>%
  # Add habitat-method column
  mutate(habitat_method = paste(environment, method, sep =" ")) %>% 
  mutate(habitat_method = case_when(
    str_detect(habitat_method, "unknown") ~ "unknown", 
    TRUE ~ habitat_method
  )) %>% 
  # Set factor levels
  mutate(habitat_method = factor(habitat_method, levels = c("marine capture", "inland capture",
                                                            "marine aquaculture", "inland aquaculture",
                                                            "unknown" )))

@@

global_export_by_source <- global_export_by_source %>%
  group_by(year, habitat_method) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  ggplot(aes(x = year, y = live_weight_t/1000000, fill = habitat_method)) +
  geom_area() +
  scale_fill_manual(values = habitat_source_colors) +
  labs(y = "Exports (mil. t, live weight)", fill = "Source", x = "") +
  theme_bw() +
  theme(legend.position = "bottom")

@@

# One attempt to include unknown - would probably need to revisit this (perhaps by country, though that would ignore the possibility of it being sourced from elsewhere)

# unknown_habitat_export<- artis %>% 
#   filter(dom_source == "error export", environment == "unknown") %>%
#   group_by(year) %>%
#   summarise(unknown_habitat_live_weight_t = sum(live_weight_t)) 
# 
# unknown_method_export<- artis %>% 
#   filter(dom_source == "error export", environment != "unknown") %>%
#   group_by(year, environment) %>%
#   summarise(unknown_method_live_weight_t = sum(live_weight_t))

# dom_export_by_source %>%
#   left_join(unknown_habitat_export, by = c("year")) %>%
#   left_join(unknown_method_export, by = c("year", "environment")) %>%
#   left_join(prod_by_source, by = c("year", "habitat_method")) %>% 
#   mutate(unknown_method_live_weight_t = ifelse(is.na(unknown_method_live_weight_t), 0, unknown_method_live_weight_t)) %>%
#   mutate(percent_export_lower = 100*dom_live_weight_t/production_t,
#          percent_export_upper = 100*(dom_live_weight_t + unknown_habitat_live_weight_t + unknown_method_live_weight_t)/production_t) %>%
#   ggplot(aes(x = year, color = habitat_method)) +
#   geom_line(aes(y = percent_export_lower)) +
#   geom_line(aes(y = percent_export_upper)) +
#   scale_color_manual(values = habitat_source_colors) +
#   labs(y = "Percent of global production exported", x = "") +
#   facet_wrap(~habitat_method) +
#   theme_bw() +
#   theme(legend.position = "none")

percent_export_by_source <- dom_export_by_source %>%
  left_join(prod_by_source, by = c("year", "habitat_method")) %>% 
  mutate(percent_export = 100*dom_live_weight_t/production_t) %>%
  ggplot(aes(x = year, y = percent_export, color = habitat_method)) +
  geom_line() +
  scale_color_manual(values = habitat_source_colors) +
  labs(y = "% production exported", x = "", color = "Source") +
  theme_bw()

@@

out_degree <- artis %>% 
  filter(habitat_method != "unknown") %>%
  select(year, exporter_iso3c, importer_iso3c, habitat_method) %>%
  distinct() %>%
  group_by(year, exporter_iso3c, habitat_method) %>%
  tally() %>%
  group_by(year, habitat_method) %>%
  ggplot(aes(x = year, y = n, color = habitat_method)) +
  geom_smooth() +
  scale_color_manual(values = habitat_source_colors) +
  labs(y = "Out degree", fill = "Source") + 
  theme_bw()

in_degree <- artis %>% 
  filter(habitat_method != "unknown") %>%
  select(year, exporter_iso3c, importer_iso3c, habitat_method) %>%
  distinct() %>%
  group_by(year, importer_iso3c, habitat_method) %>%
  tally() %>%
  group_by(year, habitat_method) %>%
  ggplot(aes(x = year, y = n, color = habitat_method)) +
  geom_smooth() +
  scale_color_manual(values = habitat_source_colors) +
  labs(y = "In degree", fill = "Source") + 
  theme_bw()

@@

```{r fig1, fig.height = 3, fig.width = 4.5}
ggarrange(global_export_by_source, 
          ggarrange(percent_export_by_source, 
                    out_degree, 
                    ncol = 2, 
                    legend = "none",
                    labels = c("b", "c")), 
          common.legend = TRUE, legend = "bottom",
          ncol = 1, labels = c("a", ""))

@@

```{r SI_top_capture_aqua_imp_exp, fig.height=4, fig.width=4}

# Summarize total imports and exports by habitat and environment
bilateral_habitat_method_summary <- artis %>%
  group_by(year, exporter_iso3c, exporter_region, importer_iso3c, importer_region, habitat_method) %>%
  summarise(live_weight_t = sum(live_weight_t))

# Top exporters, averaged over beginning and end time periods
top_exp_1 <- artis %>% 
  filter(year %in% 1996:2000) %>%
  group_by(exporter_iso3c, method) %>%
  summarise(live_weight_t = sum(live_weight_t)/length(1996:2000)) %>% 
  ungroup() %>%
  group_by(exporter_iso3c) %>%
  mutate(total = sum(live_weight_t)) %>%
  ungroup() %>%
  mutate(exporter_iso3c = fct_reorder(as.factor(exporter_iso3c), total)) %>%
  slice_max(n = 60, order_by = total) %>%
  ggplot(aes(x = live_weight_t/1000000, y = exporter_iso3c, fill = method)) +
  geom_bar(stat="identity", position = "stack") + 
  scale_fill_manual(values = method_colors) +
  lims(x = c(0, 10)) +
  labs(x = "Seafood exports (million t, live weight)", y = "",
       title = "Top 20 seafood \nexporters (1996-2000)", fill = "Production method") +
  theme_minimal()

top_exp_2 <- artis %>% 
  filter(year %in% 2015:2019) %>%
  group_by(exporter_iso3c, method) %>%
  summarise(live_weight_t = sum(live_weight_t)/length(2015:2019)) %>% 
  ungroup() %>%
  group_by(exporter_iso3c) %>%
  mutate(total = sum(live_weight_t)) %>%
  ungroup() %>%
  mutate(exporter_iso3c = fct_reorder(as.factor(exporter_iso3c), total)) %>%
  slice_max(n = 60, order_by = total) %>%
  ggplot(aes(x = live_weight_t/1000000, y = exporter_iso3c, fill = method)) +
  geom_bar(stat="identity", position = "stack") + 
  scale_fill_manual(values = method_colors) +
  lims(x = c(0, 10)) +
  labs(x = "Seafood exports (million t, live weight)", y = "",
       title = "Top 20 seafood \nexporters (2015-2019)", fill = "Production method") +
  theme_minimal()

ggarrange(top_exp_1, top_exp_2, common.legend = TRUE)

# Top importers, averaged over beginning and end time periods
top_imp_1 <- artis %>% 
  filter(year %in% 1996:2000) %>%
  group_by(importer_iso3c, method) %>%
  summarise(live_weight_t = sum(live_weight_t)/length(1996:2000)) %>% 
  ungroup() %>%
  group_by(importer_iso3c) %>%
  mutate(total = sum(live_weight_t)) %>%
  ungroup() %>%
  mutate(importer_iso3c = fct_reorder(as.factor(importer_iso3c), total)) %>%
  slice_max(n = 60, order_by = total) %>%
  ggplot(aes(x = live_weight_t/1000000, y = importer_iso3c, fill = method)) +
  geom_bar(stat="identity", position = "stack") + 
  scale_fill_manual(values = method_colors) +
  lims(x = c(0, 10)) +
  labs(x = "Seafood imports (million t, live weight)", y = "",
       title = "Top 20 seafood \nimporters (1996-2000)", fill = "Production method") +
  theme_minimal()

top_imp_2 <- artis %>% 
  filter(year %in% 2015:2019) %>%
  group_by(importer_iso3c, method) %>%
  summarise(live_weight_t = sum(live_weight_t)/length(2015:2019)) %>% 
  ungroup() %>%
  group_by(importer_iso3c) %>%
  mutate(total = sum(live_weight_t)) %>%
  ungroup() %>%
  mutate(importer_iso3c = fct_reorder(as.factor(importer_iso3c), total)) %>%
  slice_max(n = 60, order_by = total) %>%
  ggplot(aes(x = live_weight_t/1000000, y = importer_iso3c, fill = method)) +
  geom_bar(stat="identity", position = "stack") + 
  scale_fill_manual(values = method_colors) +
  lims(x = c(0, 10)) +
  labs(x = "Seafood imports (million t, live weight)", y = "",
       title = "Top 20 seafood \nimporters (2015-2019)", fill = "Production method") +
  theme_minimal()

ggarrange(top_imp_1, top_imp_2, common.legend = TRUE)

ggarrange(top_exp_2, top_imp_2, common.legend = TRUE)

# Top capture and aquaculture exporters, averaged over beginning and end time periods
top_capture_exp_1 <- artis %>% 
  filter(year %in% 1996:2000, method == "capture") %>%
  group_by(exporter_iso3c, environment) %>%
  summarise(live_weight_t = sum(live_weight_t)/length(1996:2000)) %>% 
  ungroup() %>%
  group_by(exporter_iso3c) %>%
  mutate(total = sum(live_weight_t)) %>%
  ungroup() %>%
  mutate(exporter = countrycode(exporter_iso3c, origin = "iso3c", destination = "country.name")) %>%
  mutate(exporter = fct_reorder(as.factor(exporter), total)) %>%
  slice_max(n = 2*15, order_by = total) %>%
  ggplot(aes(x = live_weight_t/1000000, y = exporter, fill = environment)) +
  geom_bar(stat="identity", position = "stack") + 
  scale_fill_manual(values = environment_colors) +
  lims(x = c(0, 4.75)) +
  labs(x = "Ave. annual exports (million t, live weight)", y = "", fill = "Production environment", 
       title = "Top 15 capture \nexporters (1996-2000)") +
  theme_minimal()

top_capture_exp_2 <- artis %>% 
  filter(year %in% 2015:2019, method == "capture") %>%
  group_by(exporter_iso3c, environment) %>%
  summarise(live_weight_t = sum(live_weight_t)/length(2015:2019)) %>% 
  ungroup() %>%
  group_by(exporter_iso3c) %>%
  mutate(total = sum(live_weight_t)) %>%
  ungroup() %>%
  mutate(exporter = countrycode(exporter_iso3c, origin = "iso3c", destination = "country.name")) %>%
  mutate(exporter = fct_reorder(as.factor(exporter), total)) %>%
  slice_max(n = 2*15, order_by = total) %>%
  ggplot(aes(x = live_weight_t/1000000, y = exporter, fill = environment)) +
  geom_bar(stat="identity", position = "stack") + 
  scale_fill_manual(values = environment_colors) +
  lims(x = c(0, 4.75)) +
  labs(x = "Ave. annual exports (million t, live weight)", y = "", fill = "Production environment",
       title = "Top 15 capture \nexporters (2015-2019)") +
  theme_minimal()

top_aqua_exp_1 <- artis %>% 
  filter(year %in% 1996:2000, method == "aquaculture") %>%
  group_by(exporter_iso3c, environment) %>%
  summarise(live_weight_t = sum(live_weight_t)/length(1996:2000)) %>% 
  ungroup() %>%
  group_by(exporter_iso3c) %>%
  mutate(total = sum(live_weight_t)) %>%
  ungroup() %>%
  mutate(exporter = countrycode(exporter_iso3c, origin = "iso3c", destination = "country.name")) %>%
  mutate(exporter = fct_reorder(as.factor(exporter), total)) %>%
  slice_max(n = 2*15, order_by = total) %>%
  ggplot(aes(x = live_weight_t/1000000, y = exporter, fill = environment)) +
  scale_fill_manual(values = environment_colors) +
  geom_bar(stat="identity", position = "stack") + 
  lims(x = c(0, 4.75)) +
  labs(x = "Ave. annual exports (million t, live weight)", y = "", fill = "Production environment",
       title = "Top 15 aquaculture \nexporters (1996-2000)") +
  theme_minimal()

top_aqua_exp_2 <- artis %>% 
  filter(year %in% 2015:2019, method == "aquaculture") %>%
  group_by(exporter_iso3c, environment) %>%
  summarise(live_weight_t = sum(live_weight_t)/length(2015:2019)) %>% 
  ungroup() %>%
  group_by(exporter_iso3c) %>%
  mutate(total = sum(live_weight_t)) %>%
  ungroup() %>%
  mutate(exporter = countrycode(exporter_iso3c, origin = "iso3c", destination = "country.name")) %>%
  mutate(exporter = fct_reorder(as.factor(exporter), total)) %>%
  slice_max(n = 2*15, order_by = total) %>%
  ggplot(aes(x = live_weight_t/1000000, y = exporter, fill = environment)) +
  scale_fill_manual(values = environment_colors) +
  geom_bar(stat="identity", position = "stack") + 
  lims(x = c(0, 4.75)) +
  labs(x = "Ave. annual exports (million t, live weight)", y = "", fill = "Production environment",
       title = "Top 15 aquaculture \nexporters (2015-2019)") +
  theme_minimal()

ggarrange(top_capture_exp_1, top_capture_exp_2, top_aqua_exp_1, top_aqua_exp_2, common.legend = TRUE, legend = "bottom")

# Top capture and aquaculture importers, averaged over beginning and end time periods
top_capture_imp_1 <- artis %>% 
  filter(year %in% 1996:2000, method == "capture") %>%
  group_by(importer_iso3c, environment) %>%
  summarise(live_weight_t = sum(live_weight_t)/length(1996:2000)) %>% 
  ungroup() %>%
  group_by(importer_iso3c) %>%
  mutate(total = sum(live_weight_t)) %>%
  ungroup() %>%
  mutate(importer = countrycode(importer_iso3c, origin = "iso3c", destination = "country.name")) %>%
  mutate(importer = fct_reorder(as.factor(importer), total)) %>%
  slice_max(n = 2*15, order_by = total) %>%
  ggplot(aes(x = live_weight_t/1000000, y = importer, fill = environment)) +
  geom_bar(stat="identity", position = "stack") + 
  scale_fill_manual(values = environment_colors) +
  lims(x = c(0, 8)) +
  labs(x = "Ave. annual imports (million t, live weight)", y = "", fill = "Production environment", 
       title = "Top 15 capture \nimporters (1996-2000)") +
  theme_minimal()

top_capture_imp_2 <- artis %>% 
  filter(year %in% 2015:2019, method == "capture") %>%
  group_by(importer_iso3c, environment) %>%
  summarise(live_weight_t = sum(live_weight_t)/length(2015:2019)) %>% 
  ungroup() %>%
  group_by(importer_iso3c) %>%
  mutate(total = sum(live_weight_t)) %>%
  ungroup() %>%
  mutate(importer = countrycode(importer_iso3c, origin = "iso3c", destination = "country.name")) %>%
  mutate(importer = fct_reorder(as.factor(importer), total)) %>%
  slice_max(n = 2*15, order_by = total) %>%
  ggplot(aes(x = live_weight_t/1000000, y = importer, fill = environment)) +
  geom_bar(stat="identity", position = "stack") + 
  scale_fill_manual(values = environment_colors) +
  lims(x = c(0, 8)) +
  labs(x = "Ave. annual imports (million t, live weight)", y = "", fill = "Production environment",
       title = "Top 15 capture \nimporters (2015-2019)") +
  theme_minimal()

>>>>

library(tidyverse)
library(countrycode)
library(circlize)
library(here)
library(ggpubr)
library(kableExtra)
library(DBI)
library(janitor)
library(viridis)
library(broom)
library(exploreARTIS)
library(vegan)
library(tidytext)
```

@@

artis <- artis %>%
  select(-record_id)

prod <- dbGetQuery(con, "SELECT * FROM production")
prod <- prod %>%
  select(-record_id)

@@

code_max_resolved_taxa <- dbGetQuery(con, "SELECT * FROM code_max_resolved_taxa") %>%
              mutate(hs6 = as.character(hs6)) %>%
  select(-record_id)

sciname_metadata <- dbGetQuery(con, "SELECT * FROM sciname") %>%
  select(-record_id)

@@

# Set directories 
outdir <- "/Volumes/jgephart/ARTIS/Outputs/ms_seafood_globalization"

@@

```{r load_drive_data}

@@

```{r join_data}
# Summarize cleaned FAO production data
prod <- prod %>% 
  select("year", "iso3c", "sciname", 
         "method", "habitat", 
         "production_t" = "live_weight_t") %>%
  group_by(year, iso3c, sciname, method, habitat) %>%
  summarise(production_t = sum(production_t, na.rm = TRUE)) %>%
  ungroup()

@@

# Update artis data to lowest resolution by comparing code taxa_level and prod taxa_level
artis <- artis %>%
  left_join(code_max_resolved_taxa,
            by = c("hs_version", "hs6", "sciname")) %>%
  left_join(sciname_metadata, 
            by = "sciname") %>%
  # Leave any missing sciname_hs_modified as original sciname
  mutate(sciname_hs_modified = case_when(
    is.na(sciname_hs_modified) ~ sciname,
    TRUE ~ sciname_hs_modified
  )) %>%
  rename(environment = habitat) %>%
  # Add habitat-method column
  mutate(habitat_method = paste(environment, method, sep =" ")) %>% 
  mutate(habitat_method = case_when(
    str_detect(habitat_method, "unknown") ~ "unknown", 
    TRUE ~ habitat_method
  )) %>% 
  # Set factor levels
  mutate(habitat_method = factor(habitat_method, levels = c("marine capture", "inland capture",
                                                            "marine aquaculture", "inland aquaculture",
                                                            "unknown" )))

@@

fig1a <- global_export_by_source %>%
  group_by(year, habitat_method) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  ggplot(aes(x = year, y = live_weight_t/1000000, fill = habitat_method)) +
  geom_area() +
  scale_fill_manual(values = habitat_source_colors) +
  labs(y = "Exports (mil. t, live weight)", fill = "Source", x = "") +
  theme_bw() +
  theme(legend.position = "bottom")

@@

percent_export_by_source <- dom_export_by_source %>%
  left_join(prod_by_source, by = c("year", "habitat_method")) %>% 
  mutate(percent_export = 100*dom_live_weight_t/production_t)

write.csv(percent_export_by_source, "outputs/percent_export_by_source_summary.csv", row.names = FALSE)

percent_export_total <- dom_export_by_source %>%
  left_join(prod_by_source, by = c("year", "habitat_method")) %>% 
  group_by(year) %>% 
  summarise(dom_live_weight_t = sum(dom_live_weight_t), 
            production_t = sum(production_t)) %>%
  mutate(percent_export = 100*dom_live_weight_t/production_t) %>% 
  mutate(calc_type = "domestic exports") %>% 
  select(year, percent_export, calc_type)

percent_export_total_all_source <- artis %>% 
  group_by(year) %>% 
  summarise(live_weight_t = sum(live_weight_t)) %>%
  left_join(prod_by_source %>% 
              group_by(year) %>% 
              summarise(production_t = sum(production_t)), 
            by = c("year")) %>% 
  mutate(percent_export = 100*live_weight_t/production_t) %>% 
  mutate(calc_type = "all export sources") %>% 
  select(year, percent_export, calc_type)

percent_export_total <- percent_export_total %>%
  bind_rows(percent_export_total_all_source)

write.csv(percent_export_total, "outputs/percent_export_total_summary.csv", row.names = FALSE)

fig1b <- ggplot() +
  geom_line(data = percent_export_by_source, aes(x = year, y = percent_export, color = habitat_method)) +
  scale_color_manual(values = habitat_source_colors) +
  labs(y = "% production exported", x = "", color = "Source") +
  theme_bw()

@@

fig1c_out_degree <- artis %>% 
  filter(habitat_method != "unknown") %>%
  select(year, exporter_iso3c, importer_iso3c, habitat_method) %>%
  distinct() %>%
  group_by(year, exporter_iso3c, habitat_method) %>%
  tally() %>%
  group_by(year, habitat_method) %>%
  ggplot(aes(x = year, y = n, color = habitat_method)) +
  geom_smooth() +
  scale_color_manual(values = habitat_source_colors) +
  labs(y = "Out degree", fill = "Source") + 
  theme_bw()

fig1c_in_degree <- artis %>% 
  filter(habitat_method != "unknown") %>%
  select(year, exporter_iso3c, importer_iso3c, habitat_method) %>%
  distinct() %>%
  group_by(year, importer_iso3c, habitat_method) %>%
  tally() %>%
  group_by(year, habitat_method) %>%
  ggplot(aes(x = year, y = n, color = habitat_method)) +
  geom_smooth() +
  scale_color_manual(values = habitat_source_colors) +
  labs(y = "In degree", fill = "Source") + 
  theme_bw()

@@

```{r fig1, fig.height = 3, fig.width = 4.5}
png("outputs/fig1.png")
ggarrange(fig1a,
          ggarrange(fig1b,
                    fig1c_out_degree,
                    ncol = 2, nrow = 1,
                    legend = "none",
                    labels = c("b", "c")),
          common.legend = TRUE, legend = "bottom",
          ncol = 1, nrow = 2, labels = c("a", ""))

dev.off()
```

```{r SI_top_capture_aqua_imp_exp, fig.height=4, fig.width=4}
# Summarize total imports and exports by habitat and environment
bilateral_habitat_method_summary <- artis %>%
  # Remove if we decide to leave FMFO in
  filter(hs6 != "230120") %>%
  group_by(year, exporter_iso3c, exporter_region, importer_iso3c, importer_region, habitat_method) %>%
  summarise(live_weight_t = sum(live_weight_t))

top_traders_recent <- bilateral_habitat_method_summary %>% 
  filter(year %in% 2016:2020, habitat_method != "unknown") %>%
  group_by(importer_iso3c, exporter_iso3c, habitat_method) %>%
  summarise(live_weight_t = sum(live_weight_t)/length(2016:2020))

top_importers_recent <- top_traders_recent %>%
  group_by(importer_iso3c, habitat_method) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  group_by(habitat_method) %>%
  slice_max(n = 10, order_by = live_weight_t) %>%
  ungroup() %>%
  group_by(habitat_method) %>%
  mutate(ranking = rank(live_weight_t)) %>%
  ungroup() %>%
  mutate(importer_iso3c = reorder_within(importer_iso3c, ranking, habitat_method)) %>%
  ggplot(aes(x = live_weight_t/1000000, y = importer_iso3c, fill = habitat_method)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = habitat_source_colors) +
  scale_y_reordered() +
  labs(y = "", x = "Live weight (million t)") +
  facet_wrap(~habitat_method, ncol = 1, scales = "free_y") +
  theme_minimal() +
  theme(legend.position = "none")

top_exporters_recent <- top_traders_recent %>%
  group_by(exporter_iso3c, habitat_method) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  group_by(habitat_method) %>%
  slice_max(n = 10, order_by = live_weight_t) %>%
  ungroup() %>%
  group_by(habitat_method) %>%
  mutate(ranking = rank(live_weight_t)) %>%
  ungroup() %>%
  mutate(exporter_iso3c = reorder_within(exporter_iso3c, ranking, habitat_method)) %>%
  ggplot(aes(x = live_weight_t/1000000, y = exporter_iso3c, fill = habitat_method)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = habitat_source_colors) +
  scale_y_reordered() +
  labs(y = "", x = "Live weight (million t)") +
  facet_wrap(~habitat_method, ncol = 1, scales = "free_y") +
  theme_minimal() +
  theme(legend.position = "none")

ggarrange(top_exporters_recent, top_importers_recent, ncol = 2, labels = c("Export", "Import"))

top_traders_past <- bilateral_habitat_method_summary %>% 
  filter(year %in% 1996:2000, habitat_method != "unknown") %>%
  group_by(importer_iso3c, exporter_iso3c, habitat_method) %>%
  summarise(live_weight_t = sum(live_weight_t)/length(1996:2000))

top_importers_past <- top_traders_past %>%
  group_by(importer_iso3c, habitat_method) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  group_by(habitat_method) %>%
  slice_max(n = 10, order_by = live_weight_t) %>%
  ungroup() %>%
  group_by(habitat_method) %>%
  mutate(ranking = rank(live_weight_t)) %>%
  ungroup() %>%
  mutate(importer_iso3c = reorder_within(importer_iso3c, ranking, habitat_method)) %>%
  ggplot(aes(x = live_weight_t/1000000, y = importer_iso3c, fill = habitat_method)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = habitat_source_colors) +
  scale_y_reordered() +
  labs(y = "", x = "Live weight (million t)") +
  facet_wrap(~habitat_method, ncol = 1, scales = "free_y") +
  theme_minimal() +
  theme(legend.position = "none")

top_exporters_past <- top_traders_past %>%
  group_by(exporter_iso3c, habitat_method) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  group_by(habitat_method) %>%
  slice_max(n = 10, order_by = live_weight_t) %>%
  ungroup() %>%
  group_by(habitat_method) %>%
  mutate(ranking = rank(live_weight_t)) %>%
  ungroup() %>%
  mutate(exporter_iso3c = reorder_within(exporter_iso3c, ranking, habitat_method)) %>%
  ggplot(aes(x = live_weight_t/1000000, y = exporter_iso3c, fill = habitat_method)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = habitat_source_colors) +
  scale_y_reordered() +
  labs(y = "", x = "Live weight (million t)") +
  facet_wrap(~habitat_method, ncol = 1, scales = "free_y") +
  theme_minimal() +
  theme(legend.position = "none")

ggarrange(top_exporters_past, top_importers_past, ncol = 2, labels = c("Export", "Import"))

# Largest export increases
top_exporter_change <- bilateral_habitat_method_summary %>% 
  filter(year %in% c(1996:2000, 2016:2020)) %>%
  mutate(year_group = case_when(
    year %in% 1996:2000 ~ "beginning", 
    year %in% 2016:2020 ~ "end")) %>%
  group_by(year_group, exporter_iso3c, habitat_method) %>% 
  summarise(live_weight_t = sum(live_weight_t)) %>%
  pivot_wider(names_from = year_group, values_from = live_weight_t) %>% 
  mutate(change = end-beginning) %>% 
  arrange(desc(change))

# Largest increases
top_export_increases <- top_exporter_change %>% 
  filter(habitat_method != "unknown") %>%
  ungroup() %>%
  group_by(habitat_method) %>%
  slice_max(n = 10, order_by = change) %>%
  ungroup() %>%
  mutate(exporter_iso3c = reorder_within(exporter_iso3c, change, habitat_method)) %>%
  ggplot(aes(x = change/1000000, y = exporter_iso3c, fill = habitat_method, color = habitat_method)) +
  geom_segment(aes(x=0, xend=change/1000000, y=exporter_iso3c, yend=exporter_iso3c)) +
  geom_point(size=4) +
  scale_color_manual(values = habitat_source_colors) +
  scale_y_reordered() +
  labs(y = "", x = "Live weight (million t)") +
  facet_wrap(~habitat_method, ncol = 1, scales = "free_y") +
  theme_minimal() +
  theme(legend.position = "none")

# Largest decreases
top_export_decreases <- top_exporter_change %>% 
  filter(habitat_method != "unknown") %>%
  ungroup() %>%
  group_by(habitat_method) %>%
  slice_min(n = 10, order_by = change) %>%
  ungroup() %>%
  mutate(exporter_iso3c = reorder_within(exporter_iso3c, change, habitat_method)) %>%
  ggplot(aes(x = change/1000000, y = exporter_iso3c, fill = habitat_method, color = habitat_method)) +
  geom_segment(aes(x=0, xend=change/1000000, y=exporter_iso3c, yend=exporter_iso3c)) +
  geom_point(size=4) +
  scale_color_manual(values = habitat_source_colors) +
  scale_y_reordered() +
  labs(y = "", x = "Live weight (million t)") +
  facet_wrap(~habitat_method, ncol = 1, scales = "free_y") +
  theme_minimal() +
  theme(legend.position = "none")

ggarrange(top_export_decreases, top_export_increases, nrow = 1)

# Largest import increases
top_importer_change <- bilateral_habitat_method_summary %>% 
  filter(year %in% c(1996:2000, 2016:2020)) %>%
  mutate(year_group = case_when(
    year %in% 1996:2000 ~ "beginning", 
    year %in% 2016:2020 ~ "end")) %>%
  group_by(year_group, importer_iso3c, habitat_method) %>% 
  summarise(live_weight_t = sum(live_weight_t)) %>%
  pivot_wider(names_from = year_group, values_from = live_weight_t) %>% 
  mutate(change = end-beginning) %>% 
  arrange(desc(change))

# Largest increases
top_import_increases <-top_importer_change %>% 
  filter(habitat_method != "unknown") %>%
  ungroup() %>%
  group_by(habitat_method) %>%
  slice_max(n = 10, order_by = change) %>%
  ungroup() %>%
  mutate(importer_iso3c = reorder_within(importer_iso3c, change, habitat_method)) %>%
  ggplot(aes(x = change/1000000, y = importer_iso3c, fill = habitat_method, color = habitat_method)) +
  geom_segment(aes(x=0, xend=change/1000000, y=importer_iso3c, yend=importer_iso3c)) +
  geom_point(size=4) +
  scale_color_manual(values = habitat_source_colors) +
  scale_y_reordered() +
  labs(y = "", x = "Live weight (million t)") +
  facet_wrap(~habitat_method, ncol = 1, scales = "free_y") +
  theme_minimal() +
  theme(legend.position = "none")

# Largest decreases
top_import_decreases <-top_importer_change %>% 
  filter(habitat_method != "unknown") %>%
  ungroup() %>%
  group_by(habitat_method) %>%
  slice_min(n = 10, order_by = change) %>%
  ungroup() %>%
  mutate(importer_iso3c = reorder_within(importer_iso3c, change, habitat_method)) %>%
  ggplot(aes(x = change/1000000, y = importer_iso3c, fill = habitat_method, color = habitat_method)) +
  geom_segment(aes(x=0, xend=change/1000000, y=importer_iso3c, yend=importer_iso3c)) +
  geom_point(size=4) +
  scale_color_manual(values = habitat_source_colors) +
  scale_y_reordered() +
  labs(y = "", x = "Live weight (million t)") +
  facet_wrap(~habitat_method, ncol = 1, scales = "free_y") +
  theme_minimal() +
  theme(legend.position = "none")

ggarrange(top_import_decreases, top_import_increases, nrow = 1)

# Repeat for regional trends
# Largest export increases
top_exporter_change <- bilateral_habitat_method_summary %>% 
  filter(year %in% c(1996:2000, 2016:2020)) %>%
  mutate(year_group = case_when(
    year %in% 1996:2000 ~ "beginning", 
    year %in% 2016:2020 ~ "end")) %>%
  group_by(year_group, exporter_region, habitat_method) %>% 
  summarise(live_weight_t = sum(live_weight_t)) %>%
  pivot_wider(names_from = year_group, values_from = live_weight_t) %>% 
  mutate(change = end-beginning) %>% 
  arrange(desc(change))

# Largest increases
top_region_export <- top_exporter_change %>% 
  filter(habitat_method != "unknown") %>%
  ungroup() %>%
  group_by(habitat_method) %>%
  slice_max(n = 10, order_by = change) %>%
  ungroup() %>%
  mutate(exporter_region = reorder_within(exporter_region, change, habitat_method)) %>%
  ggplot(aes(x = change/1000000, y = exporter_region, fill = habitat_method, color = habitat_method)) +
  geom_segment(aes(x=0, xend=change/1000000, y=exporter_region, yend=exporter_region)) +
  geom_point(size=4) +
  scale_color_manual(values = habitat_source_colors) +
  scale_y_reordered() +
  labs(y = "", x = "Live weight (million t)") +
  facet_wrap(~habitat_method, ncol = 1, scales = "free_y") +
  theme_minimal() +
  theme(legend.position = "none")

# Largest import increases
top_importer_change <- bilateral_habitat_method_summary %>% 
  filter(year %in% c(1996:2000, 2016:2020)) %>%
  mutate(year_group = case_when(
    year %in% 1996:2000 ~ "beginning", 
    year %in% 2016:2020 ~ "end")) %>%
  group_by(year_group, importer_region, habitat_method) %>% 
  summarise(live_weight_t = sum(live_weight_t)) %>%
  pivot_wider(names_from = year_group, values_from = live_weight_t) %>% 
  mutate(change = end-beginning) %>% 
  arrange(desc(change))

# Largest increases
top_region_import <-top_importer_change %>% 
  filter(habitat_method != "unknown") %>%
  ungroup() %>%
  group_by(habitat_method) %>%
  slice_max(n = 10, order_by = change) %>%
  ungroup() %>%
  mutate(importer_region = reorder_within(importer_region, change, habitat_method)) %>%
  ggplot(aes(x = change/1000000, y = importer_region, fill = habitat_method, color = habitat_method)) +
  geom_segment(aes(x=0, xend=change/1000000, y=importer_region, yend=importer_region)) +
  geom_point(size=4) +
  scale_color_manual(values = habitat_source_colors) +
  scale_y_reordered() +
  labs(y = "", x = "Live weight (million t)") +
  facet_wrap(~habitat_method, ncol = 1, scales = "free_y") +
  theme_minimal() +
  theme(legend.position = "none")

ggarrange(top_region_export, top_region_import, nrow = 1, labels = c("Export", "Import"))

@@
